---
title: "Class 4"
author: "Colin Kuehl"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: pdf_document
editor_options: 
  chunk_output_type: console
---
```{r,echo=FALSE,message=FALSE,warning=FALSE}
library(formatR)#may need to install this first
# Set so that long lines in R will be wrapped:
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=80), tidy=TRUE)
```

# We'll start out illustrating concepts from discussion using R


## Distribution of the Mean     
```{r}
getmean=function(N, mean, sd){
  opinion=rnorm(n = N, mean=mean,sd=sd)
  meanopinion=mean(opinion)
  return(meanopinion)} 
#Building a function in R. This is beyond the scope of the class

manymeansa=replicate(n = 10, getmean(20, mean=65, sd=6))
hist(manymeansa, main="Many Means 10 Draws", sub=mean(manymeansa))
mean(manymeansa)
abline(v=mean(manymeansa), col="orange")

#look with different n
manymeansb=replicate(n = 20, getmean(20, mean=65, sd=6))
hist(manymeansb, main="Many Means 20 Draws", sub=mean(manymeansb))
mean(manymeansb)
abline(v=mean(manymeansb), col="orange")


manymeansc=replicate(n = 100, getmean(20, mean=65, sd=6))
hist(manymeansc, main="Many Means 100 Draws", sub=mean(manymeansc))
mean(manymeansc)
abline(v=mean(manymeansc), col="orange")


manymeansd=replicate(n = 1000, getmean(20, mean=65, sd=6))
hist(manymeansd, main="Many Means 1k Draws", sub=mean(manymeansd))
mean(manymeansd)
abline(v=mean(manymeansd), col="orange")

manymeanse=replicate(n = 1000, getmean(20, mean=65, sd=6))
hist(manymeanse, main="Many Means 1000 Draws", sub=mean(manymeanse))
mean(manymeanse)
abline(v=mean(manymeanse), col="orange")

manymeansf=replicate(n = 10000, getmean(20, mean=65, sd=6))
hist(manymeansf, main="Many Means 10k Draws", sub=mean(manymeansf))
mean(manymeansf)
abline(v=mean(manymeansf), col="orange")

#Larger sample sizes should be closer to "true mean" but still variation
abline(v=65, col="blue")
abline(v=mean(manymeansa), col="darkorange4")
abline(v=mean(manymeansb), col="darkorange3")
abline(v=mean(manymeansc), col="darkorange2")
abline(v=mean(manymeansd), col="darkorange1")
abline(v=mean(manymeanse), col="darkorange")
abline(v=65, col="blue")



mean(manymeansa)
mean(manymeansb)
mean(manymeansc)
mean(manymeansd)
mean(manymeanse)
mean(manymeansf) #as n increases will get closer to true mean

sd(manymeansa) # . . . and sd
sd(manymeansb)
sd(manymeansc)
sd(manymeansd)
sd(manymeanse)
sd(manymeansf)
```
     
           
        



We have two datasets today. The first looks at a measure of the concept of idealogy in congress. The second is about gender and the movie industry. 

As always Set your working and then Load the data(congress)
```{r}
setwd("~/Dropbox/POLS 641/2025 641/Rcode F25/Week 4")#you will need to change for your folders
load("congress.RData")
```

## We'll start by exploring the validity and reliability of the DW-Nominate Scale

This is a measure of members of congress ideology

- A Wikipedia entry is available [here.](https://en.wikipedia.org/wiki/NOMINATE_(scaling_method)))
- A longer explainer is available [here.](https://www.swarthmore.edu/SocSci/rvalell1/documents/Nominate_000.pdf)
- What do you think? (in terms of validity? reliability?)


Explore the data
```{r}
head(congress)
length(unique(congress$name))

```



```{r Party Subset, echo=TRUE}
## subset the data by party
rep <- subset(congress, party == "Republican") #this command uses dplyr/tidyverse
dem <- congress[congress$party == "Democrat", ] #Uses base R, but does the same thing
```

Subset Further by Congress
```{r Congress Subset, echo=TRUE}
rep80 <- subset(rep, subset = (congress == 80))
dem80 <- subset(dem, subset = (congress == 80))
rep112 <- subset(rep, subset = (congress == 112))
dem112 <- subset(dem, subset = (congress == 112))
```

## Make a Scatterplot
```{r 80th Congress 1, echo=T}
plot(dem80$dwnom1, dem80$dwnom2)
```

##Make it Prettier
```{r 80th Congress 2, echo=T}
Demplot<-plot(dem80$dwnom1, dem80$dwnom2, pch = 16, col = "blue", 
     xlim = c(-1.5,1.5), ylim = c(-1.5,1.5),
     xlab = "Economic liberalism/conservatism",
     ylab = "Racial liberalism/conservatism",
     main = "80th Congress")
```
What does this tell us about the two dimensions?


##Let's add the GOP
```{r add GOP code, eval = F, echo=T}
plot(dem80$dwnom1, dem80$dwnom2, pch = 16, col = "blue", 
     xlim = c(-1.5,1.5), ylim = c(-1.5,1.5),
     xlab = "Economic liberalism/conservatism", 
     ylab = "Racial liberalism/conservatism",
     main = "80th Congress")
points(rep80$dwnom1, rep80$dwnom2, pch = 17, col = "red")
text(-0.75, 1, "Democrats") #adding a label to our plot
text(1, -1, "Republicans")
```

*When plotting, you have to send ALL the commands at once*

##Ain't it Perty?!?
```{r 80th Congress 3, echo=F}
plot(dem80$dwnom1, dem80$dwnom2, pch = 16, col = "blue", xlim = c(-1.5,1.5), ylim = c(-1.5,1.5),xlab = "Economic liberalism/conservatism", ylab = "Racial liberalism/conservatism", main = "80th Congress")
points(rep80$dwnom1, rep80$dwnom2, pch = 17, col = "red")
text(-0.75, 1, "Democrats")
text(1, -1, "Republicans")
```

##Class Exercise

*Can you produce an identical plot for the 112th Congress?*
```{r, echo=F}


```

*What do these plots show us? Does the measure have face validity?*



```{r 112th Congress, eval=T, echo=F}

plot(dem112$dwnom1, dem112$dwnom2, pch = 16, col = "blue", xlim = c(-1.5,1.5), ylim = c(-1.5,1.5),xlab = "Economic liber alism/conservatism", ylab = "Racial liberalism/conservatism", main = "112th Congress")
points(rep112$dwnom1, rep112$dwnom2, pch = 17, col = "red")
text(-0.75, 1, "Democrats")
text(1, -1, "Republicans")
```

##Let's Dig into this Further

How have the median R and D Members of Congress changed over time?

First calculate the median for each party by Congress:
```{r Medians, echo=T}
dem.median <- tapply(dem$dwnom1, dem$congress, median) #Find the median of dwnom1 for each congress. You won't need to use tapply on problem sets etc, but it is pretty handy
rep.median <- tapply(rep$dwnom1, rep$congress, median)
```
#tapply is a useful command, but you won't need it for problem sets etc. Dplyr group_by does similar things


Now let's plot those medians for each party over time:
```{r Time Series Code, eval=F, echo=T}
plot(names(dem.median), dem.median, col = "blue", type = "l",
     xlim = c(80, 115), ylim = c(-1, 1), xlab = "Congress",
     ylab = "DW-NOMINATE score (1st dimension)")
lines(names(rep.median), rep.median, col = "red")
text(110, -0.6, "Democratic\n Party")
text(110, 0.85, "Republican\n Party")
```

#Now repeat for the 2nd Dimension
```{r}
dem.median2 <- tapply(dem$dwnom2, dem$congress, median) #Find the median of dwnom1 for each congress. You won't need to use tapply on problem sets etc, but it is pretty handy
rep.median2 <- tapply(rep$dwnom2, rep$congress, median)

plot(names(dem.median2), dem.median2, col = "blue", type = "l",
     xlim = c(80, 115), ylim = c(-1, 1), xlab = "Congress",
     ylab = "DW-NOMINATE score (2nd dimension)")
lines(names(rep.median2), rep.median2, col = "red")
text(92, 0.6, "Democratic\n Party")
text(92, -0.85, "Republican\n Party")

```
Bonus: Why might we want to use the median instead of the mean here? Does using the mean change the broader story?


Really interesting viewing of DW Nominate data: https://www.youtube.com/watch?v=M193ZIOFk0k


#################################################################################################




Load the data: These data comes from this [fivethirtyeight article](https://fivethirtyeight.com/features/the-dollar-and-cents-case-against-hollywoods-exclusion-of-women/)

```{r}
load("bech.RData")
```




Load Packages: we'll use ggplot briefly again today
```{r}
library(ggplot2)
```

Find the number of variables and observations - *What is our unit of analysis?*
```{R}
head(bech)
```

VARIABLE NAMES 
```{r}
names(bech) 
variable.names(bech) 
```

What time range does the data cover?
```{r}
min(bech$year)
max(bech$year)
range(bech$year)
```

Exploring the data
```{r}
table(bech$clean_test)

plot(bech$domgross, bech$intgross)

prop.table(table(bech$binary))

prop.table(table(bech$clean_test, bech$binary))

barplot(table(bech$clean_test)) #show ugly bar plot

```

FIND THE RANGE OF GROSS REVENUE(we'll just look at domestic and standardize to 2013 dollar value)
```{r}
range(bech$domgross_2013)
max(bech$domgross_2013) #Bonus: What movie is this? Hint: View(yourdata)
min(bech$domgross_2013) #Cheers to you if you've actually watched this
```

FIND STANDARD DEVIATION AND VARIANCE - We'll take a closer look at this next week.
Formula for Standard Deviation:
$$\sigma = \sqrt{\frac{\sum\limits_{i=1}^{n} \left(x_{i} - \bar{x}\right)^{2}} {n-1}}$$
More info on using math symbols in Rmarkdown is here: https://rpruim.github.io/s341/S19/from-class/MathinRmd.html

```{r}
sqrt(sum((bech$domgross_2013-mean(bech$domgross_2013))^2/(length(bech$domgross_2013)-1))) #How would we turn this into the variance?
```

Of course, it is much easier to just have R do it
```{r}
sd(bech$domgross_2013)
var(bech$domgross_2013)
sd(bech$domgross_2013)^2 #Variance = standard deviation squared.
sqrt(var(bech$domgross_2013)) #Standard deviation = square root of variance.
```



TAKE A LOOK AT THE DISTRIBUTION OF DOMESTIC GROSS  
```{r}
hist(bech$domgross_2013, breaks = 15) #breaks=  number of bars
hist(bech$domgross_2013, main="Distribution of Gross Revenue", xlab="Gross Revenue", col="blue",border="green",breaks=30, las=2) #bedazzled
```

LOOK AT THE DENSITY PLOT OF DOMESTIC GROSS
```{r}
plot(density(bech$domgross_2013),main="Density Distribution of Gross Revenue")

ggplot(bech, aes(x=domgross_2013))+geom_density()+
  labs(x="Domestic Gross 2013", y="Density", title="Density Distribution of Gross Revenue", caption="ggplot version")
```

How are histogram and density plot different? Which is closer to a pdf?
```{r}
hist(bech$domgross_2013, main="Histogram of of Gross Revenue",
xlab="Domestic Gross 2013",breaks=20)

ggplot(bech, aes(x=domgross_2013))+geom_histogram()+
  labs(x="Domestic Gross 2013", y="Density", title="Histogram of Gross Revenue", caption="ggplot version")
```

MAKE A BOXPLOT
```{r}
boxplot(bech$domgross_2013, main="Boxplot of Gross Revenue", col="cornflowerblue")
abline(h=mean(bech$domgross_2013), col="deeppink", lwd=2)
```
A boxplot shows the median, 25% and 75% quartile, and important outliers

ADD LINES TO INDICATE THE MEAN AND MEDIAN, AND MAKE A LEGEND *(Make sure all the code stays together)*
```{r}
plot(density(bech$domgross_2013), main="Density Plot of Gross Revenue", xlab="Gross")
abline(v=mean(bech$domgross_2013), lty=2, lwd=2, col="red") # put a vertical line at the mean 
abline(v=median(bech$domgross_2013), lty=4, lwd=2, col="blue")
legend("topright", legend=c("mean", "median"), lty=c(2,4), col=c("red","blue"))

options(scipen=10)#This stops use of scientific notation
```


Take Random Samples and see how mean and median changes
```{r}
sample1 <- bech[sample(nrow(bech), 100), ] #Randomly samples 100 movies

plot(density(sample1$domgross_2013), main="Density Plot of Gross Revenue", xlab="Gross")
abline(v=mean(sample1$domgross_2013), lty=2, lwd=2, col="red") 
abline(v=median(sample1$domgross_2013), lty=4, lwd=2, col="blue")
legend("topright", legend=c(mean(sample1$domgross_2013), median(sample1$domgross_2013)), lty=c(2,4), col=c("red","blue"))


sample2 <- bech[sample(nrow(bech), 100), ]
plot(density(sample2$domgross_2013), main="Density Plot of Gross Revenue", xlab="Gross")
abline(v=mean(sample2$domgross_2013), lty=2, lwd=2, col="red") 
abline(v=median(sample2$domgross_2013), lty=4, lwd=2, col="blue")
legend("topright", legend=c(mean(sample2$domgross_2013), median(sample2$domgross_2013)), lty=c(2,4), col=c("red","blue"))


mean(sample1$domgross_2013)
mean(sample2$domgross_2013)

median(sample1$domgross_2013)
median(sample2$domgross_2013)
```



LOOK AT THE SAMPLING DISTRIBUTION OF REVENUE AND BUDGET *(Let's put the four figures arranged in 2 rows and 2 columns)*
```{r}
par(mfrow=c(2,2)) ###This line of code tells R you want a 2 x 2 of smaller graphs
hist(bech$domgross_2013, main="Histogram of Domestic Gross", breaks = 15)
plot(density(bech$domgross_2013), main="Distribution of Domestic Gross")
hist(bech$budget_2013, main="Histogram of Budget", breaks = 15)
plot(density(bech$budget_2013), main="Distribution of Budget")
```


THIS IS NOT A NORMAL DISTRIBUTION SO MAY WANT TO CREATE LOGGED VERSION
We will talk about transforming variables later
```{r}
par(mfrow=c(1,1)) #Return to just looking at one graph at a time
bech$logdomgross13 <- log(bech$domgross_2013)#creates new logged variable
hist(bech$logdomgross13, main="Logged Domestic Revenue")#Looks more normal, harder to interpret
boxplot(bech$logdomgross13, main="Logged Domestic Revenue")

hist(bech$domgross_2013, main="Domestic Revenue")#Looks more normal, harder to interpret
boxplot(bech$logdomgross13, main="Logged Domestic Revenue")

```



LET'S USE STATS TO MAKE COMPARISONS

First, find averages for films that pass and fail the Bechdel test (there are several ways to do this).
```{r}
pass <- bech[which(bech$binary=="PASS"), ] #Creates a new dataset of only those that pass the bechdel test
fail <- bech[which(bech$binary=="FAIL"), ]
mean(pass$domgross_2013)
mean(fail$domgross_2013)
#How many films passed and how many failed?

mean(pass$budget_2013)
mean(fail$budget_2013)

ggplot(bech, aes(x=logdomgross13, fill=binary)) +
    geom_density( alpha=.5, position="identity") #overlay both distributions
```

Second, simple algebra to find difference.
```{r}
grossdiff <- mean(pass$domgross_2013)-mean(fail$domgross_2013)
grossdiff
```

*Another method for finding mean of subset (more here: http://www.statmethods.net/management/subset.html)*
```{r}
xbar_pass_t <- mean(bech$domgross_2013[bech$binary=="PASS"])
xbar_fail_t <- mean(bech$domgross_2013[bech$binary=="FAIL"])
DIFF <- (xbar_pass_t - xbar_fail_t)
DIFF
```

Show Both Box Plots
```{r}
boxplot(pass$domgross_2013)
boxplot(fail$domgross_2013)
#or from original dataset
boxplot(bech$domgross_2013~bech$binary) 
#using ggplot
ggplot(bech, aes(binary,domgross_2013))+geom_boxplot()+
  labs(y="Domestic Gross 2013", x="Bechdel Test") #creates boxplots based on values of binary

#to illustrate we can use this for variety of categories(and logged)
ggplot(bech, aes(test,logdomgross13))+geom_boxplot()+
  labs(y="Domestic Gross 2013", x="Bechdel Test")+ theme(axis.text.x=element_text(angle = -45, hjust = 0))



ggplot(bech, aes(as.factor(year),logdomgross13))+geom_boxplot()+
  labs(y="Domestic Gross 2013", x="Bechdel Test")+ theme(axis.text.x=element_text(angle = -90, hjust = 0))

```

FINDING RELATIONSHIPS *(Let's see if bigger budgets are related to a higher gross)*
```{r}
par(mfrow=c(1,1)) #show one graph at a time
plot(bech$budget, bech$domgross_2013, main="Gross Revenue and Budget", xlab = "Budget", ylab = "Gross Revenue")
```


How would we create the same plot using ggplot? (hint: look at last week's code)
```{r}

```


IT'S HARD TO SEE A PATTERN FROM THE PLOT, SO LET'S MAKE A LINEAR REGRESSION MODEL: REMEMBER THE MODEL IS lm(Y ~ X, data=nameofdata)
```{r}
plot(bech$budget, bech$domgross_2013, main="Gross Revenue and Budget", xlab = "Budget", ylab = "Gross Revenue")
model1 <- lm(domgross_2013~budget, data=bech) #Making an object for a linear regression model. Notice how I access the variables used... 
summary(model1)
abline(model1, col = "purple", lty=2) #This puts the trend line

abline(lm(pass$domgross_2013~pass$budget_2013), lty=3, lwd=2, col="pink") #for fun add lines for films that pass bechdel test
abline(lm(fail$domgross_2013~fail$budget_2013), lty=4, lwd=2, col="blue")
legend("topright", legend=c("All", "Pass", "Fail"), lty=c(2,3,4), col=c("purple","pink", "blue"))
```
What do the different slopes mean?What does this tell us about the profitability of inclusing in the movie industry?


Another Subset: We may also want to exclude outliers
```{r}
middle <- subset(bech, budget_2013 >= 1000000 & budget_2013 < 10000000) #only budgets between 1 and 10 millsion
summary(middle$budget_2013)

#do the relationships change when we control for outliers(hint: use copy and paste plus find and replace)



#What about when we use the log of gross

```
################################################################

