---
title: "Research Title Here"
subtitle: "Subtitle"
author: "Your Name"
institute: "Northern Illinois University"
date: today
format:
  revealjs:
    theme: simple
    slide-number: true
    progress: true
    transition: fade
    footer: "Author (Year) | Policy Brief "
execute:
  echo: false
  warning: false
  message: false
---

```{r}
#| label: setup

library(tidyverse)
library(knitr)
library(kableExtra)#for tables
library(modelsummary)

# ── This makes up data that is then used throughout the template. Delete this and all other references for your actual presentation ──────────────────────────────────────────────────────────
set.seed(42)
n <- 200

df <- tibble(
  country     = paste0("Country_", 1:n),
  gdp_pc      = exp(rnorm(n, 9, 1)),
  democracy   = rbinom(n, 1, 0.5),
  trade_open  = runif(n, 10, 100),
  outcome     = 2 + 0.3 * log(gdp_pc) + 0.8 * democracy +
                0.01 * trade_open + rnorm(n)
)
```

## Title Slide {.center}

------------------------------------------------------------------------

## Research Puzzle

::::: columns
::: {.column width="55%"}
**Motivating Question**

This is a bulleted list - First bullet - Second bullet - **Third Bullet in Bold**
:::

::: {.column width="45%"}
:::
:::::

------------------------------------------------------------------------

## Background and Literature

**What we know**

**What remains unresolved**

- Gap 1 — brief description
- Gap 2 — brief description

**This paper's answers this question**

::: notes
Speaker notes: These don't show up on the screen, but can be useful when building
:::

------------------------------------------------------------------------

## Hypotheses

Based on \[theory / mechanism\], I derive the following expectations:

::: incremental
- **H1 (Main):** As *X* increases, *Y* increases, because \[mechanism\].

- **Null Hypothesis** If \[rival theory\] holds, we would instead expect \[alternative pattern\].
:::

------------------------------------------------------------------------

## Measurement

## Descriptive Statistics

```{r}
#| label: tbl-descriptives
#| tbl-cap: "Summary Statistics"

desc_vars <- df |>
  select(
    `GDP per Capita`  = gdp_pc,
    `Democracy`       = democracy,
    `Trade Openness`  = trade_open,
    `Outcome`         = outcome
  )

datasummary_skim(
  desc_vars,
  output = "kableExtra",
  fmt    = 2
) |>
  kable_styling(
    font_size         = 20,
    bootstrap_options = c("striped", "hover"),
    full_width        = FALSE
  )
```

------------------------------------------------------------------------

## Descriptive Statistics (cont.)

```{r}
#| label: fig-distribution
#| fig-cap: "Distribution of Outcome by Regime Type"
#| fig-height: 4.5

df |>
  mutate(Regime = if_else(democracy == 1, "Democracy", "Autocracy")) |>
  ggplot(aes(x = outcome, fill = Regime)) +
  geom_density(alpha = 0.5, colour = NA) +
  scale_fill_manual(values = c("Democracy" = "#2166ac",
                               "Autocracy"  = "#d6604d")) +
  labs(
    x    = "Outcome",
    y    = "Density",
    fill = NULL
  ) +
  theme_minimal(base_size = 16) +
  theme(legend.position = "top")
```

------------------------------------------------------------------------

## Primary Regression Results

```{r}
#| label: tbl-main-results
#| tbl-cap: "OLS Regression — Main Results"

m1 <- lm(outcome ~ democracy,                             data = df)
m2 <- lm(outcome ~ democracy + log(gdp_pc),               data = df)
m3 <- lm(outcome ~ democracy + log(gdp_pc) + trade_open,  data = df)

modelsummary(
  list("Bivariate" = m1, "+ Controls" = m2, "Full Model" = m3),
  stars       = c("*" = .1, "**" = .05, "***" = .01),
  gof_map     = c("nobs", "r.squared", "adj.r.squared"),
  coef_rename = c(
    "democracy"    = "Democracy",
    "log(gdp_pc)"  = "ln GDP per Capita",
    "trade_open"   = "Trade Openness"
  ),
  output      = "kableExtra",
  fmt         = 3
) |>
  kable_styling(font_size = 18, full_width = FALSE) |>
  row_spec(0, bold = TRUE)
```

::: fragment
**Key finding:** Democracy is positively and significantly associated with the outcome across all specifications.
:::

------------------------------------------------------------------------

## Regression Robustness Checks

```{r}
#| label: fig-coefplot
#| fig-cap: "Coefficient Plot — Robustness Across Specifications"
#| fig-height: 4

models <- list(
  "Main Sample"  = m3,
  "Subsample A"  = lm(outcome ~ democracy + log(gdp_pc) + trade_open,
                       data = filter(df, gdp_pc < median(gdp_pc))),
  "Subsample B"  = lm(outcome ~ democracy + log(gdp_pc) + trade_open,
                       data = filter(df, gdp_pc >= median(gdp_pc)))
)

modelplot(
  models,
  coef_map   = c("democracy" = "Democracy"),
  conf_level = 0.95
) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "grey50") +
  scale_colour_brewer(palette = "Set1") +
  labs(x = "Coefficient Estimate (95% CI)", y = NULL, colour = NULL) +
  theme_minimal(base_size = 15) +
  theme(legend.position = "right")
```

The positive effect of democracy is stable across sample splits and alternative specifications.

------------------------------------------------------------------------

## Contributions & Qualitative Research Proposal

**Theoretical Contributions**

- Contribution 1 — how this paper advances existing theory
- Contribution 2 — novel mechanism or scope condition identified

**Empirical Contributions**

- Contribution 3 — new data, identification strategy, or case coverage

**Qualitative Follow-Up**

To probe the mechanism, future work should conduct \[case studies / interviews / process tracing\] in:

| Case      | Rationale                           |
|-----------|-------------------------------------|
| Country A | Most-likely case for H1             |
| Country B | Least-likely case — hard test       |
| Country C | Deviant case — unexplained by model |

------------------------------------------------------------------------

## Discussion

**Implications**

Briefly state what these findings mean for theory and / or policy.

**Limitations and future research**

- Limitation 1
- Limitation 2

## Conclusion

**Summary of findings**

- Finding 1 — one sentence
- Finding 2 — one sentence
- Finding 3 — one sentence

------------------------------------------------------------------------

##  {.center}

### Thank You

**\[Your Name\]** \[your.email\@niu.edu\]

*Slides and replication materials available at \[link\]*

::: {.callout-tip appearance="minimal"}
Questions and comments welcome
:::
